---
title: "R Notebook"
output:
editor_options:
  chunk_output_type: console
---

In this practical, you will visualize the development of wealth in Basel's quarters across time and learn how to make the visualization pretty.   

# 0 - Preliminaries

1. If you haven't done so already, load the `tidyverse` and the `taxation.csv` data. 

```{r}
library(tidyverse)
basel <- read_csv('1_data/taxation.csv')
```

# 1 - Plot base

1. Begin by creating a plot showing as lines (`geom_line()`) the relationship of `wealth_median` (`y`) and `years` (`x`) for each of the `quarter`s (`col`).   

```{r}
basel %>% 
  ggplot(aes(x = XX,
             y = XX,
             col = XX)) + 
  XX
```

# 2 - Themes

1. Make the plot prettier by adding a `theme_*()`. I like `theme_minimal()`, but go for a different one if you like, e.g., `theme_light()`. 

```{r}
basel %>% 
  ggplot(aes(x = XX,
             y = XX,
             col = XX)) + 
  XX +
  XX
```

# 3 - Theme

1. Now use `theme()` to fix the legend. I bet on your screen the legend is also taking up most of the space. Move the legend to the bottom using `legend.position = "bottom"`. 

```{r}
basel %>% 
  ggplot(aes(x = XX,
             y = XX,
             col = XX)) + 
  XX +
  XX + 
  XX(XX = XX)
```

2. That's better. If for you, the legend is now not fitting into the plotting area, use  `legend.title = element_blank()` and `legend.text = element_text(size = 7)` to make it fit as done in the presentation. 
   
```{r}
basel %>% 
  ggplot(aes(x = XX,
             y = XX,
             col = XX)) + 
  XX +
  XX + 
  XX(XX = XX,
     XX = XX,
     XX = XX)
```

# 4 - Labs

1. Ok, the plot itself is pretty decent already. What's still missing, however, is appropriate annotion. Use `labs()` to add better titles for the axes, a plot title and subtitle, and a caption.  

```{r}
basel %>% 
  ggplot(aes(x = XX,
             y = XX,
             col = XX)) + 
  XX +
  XX + 
  XX(XX = XX,
     XX = XX,
     XX = XX) +
  XX(XX = "XX",
     XX = "XX",
     XX = "XX",
     XX = "XX",
     XX = "XX")
```

# 5 - Colors

1. Add `scale_color_viridis_d` to change the color set used to the popular viridis color set. Note the `d` at the end. This refers to `discrete` as opposed to `continuous`, which is an distinction made for most scaling functions. The rule is, use `discrete` whenever you are scaling a categorical variable and `continuous` whenever you are scaling a continuous variable. Here we are a categorical variable, namely `quarter`. 

```{r}
basel %>% 
  ggplot(aes(x = XX,
             y = XX,
             col = XX)) + 
  XX +
  XX + 
  XX(XX = XX,
     XX = XX,
     XX = XX) +
  XX(XX = "XX",
     XX = "XX",
     XX = "XX",
     XX = "XX",
     XX = "XX") + 
  XX
```

# 6 - Factors

1. To change the order of categorical variables, for instance, to change the order of colors assigned to its values, you must use `factors`. The two data wrangling lines added below make sure that the factor levels in `quarter` are ordered according to `wealth_median` in year 2001. The `as_factor` function btw is part of the `forcats` package, which exists exactly to solve these kinds of problems.   

```{r}
basel %>% 
  arrange(year, wealth_median) %>% 
  mutate(quarter = as_factor(quarter)) %>% 
  ggplot(aes(x = XX,
             y = XX,
             col = XX)) + 
  XX +
  XX + 
  XX(XX = XX,
     XX = XX,
     XX = XX) +
  XX(XX = "XX",
     XX = "XX",
     XX = "XX",
     XX = "XX",
     XX = "XX") + 
  XX
```


# 6 - Go explore

1. Install the `ggthemes` package and try out some of its themes. 




